---
title: "Project 1"
output: html_document
---

***Gehad Gad***

***February 23, 2020***

***DATA 607 - Project 1***



In this project, you’re given a text file with chess tournament results where the information has some structure. Your
job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database)
with the following information for all of the players:
Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents
For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605
1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and
dividing by the total number of games played.


```{r}
#Import libraries and/or Packages

library(stringr)
library (ggplot2)
```


```{r}

# I Loaded the text file to my github to be read in r.

file = "https://github.com/GehadGad/DATA607-Project-1/raw/master/tournamentinfo.txt"

# The whole data can be seen from the link below.
#Data <- read.table (url("https://github.com/GehadGad/DATA607-Project-1/raw/master/tournamentinfo.txt"), sep = ",")

```


```{r}
#Display the head of the Data

data=readLines(file , warn=FALSE )

head(data)
```


```{r}
# Data cleaning, removing separators
data<-data[!str_detect(data,"\\-----")]

# Removing headers and separators
data<-str_split(data[3:length(data)],"\\|")

```


```{r}

#Extract names of all players
extracted<-str_extract_all(data,"[[:alpha:]-?[:alpha:]  ?]{2,}")
names<-unlist(extracted)[str_detect(unlist(extracted),"[[:alpha:]]{3,}")]
head(names)

```


```{r}
#Extract state of origin
state<-unlist(extracted)[(str_detect(unlist(extracted),"[[:alpha:]]{2}"))&(!str_detect(unlist(extracted),"[[:alpha:]]{3,}"))]
head(state)

```


```{r}
#Extract points
num_data<-str_extract_all(data,"\\d{1,}+\\.?.?")
pts<-unlist(num_data)[(str_detect(unlist(num_data),"\\d\\.\\d"))]
head(pts)
```

```{r}
#Extract pre->post
rtg<-str_extract_all(data,"(( \\:)|(\\>))?.?\\d{1,}P*\\.?")
pre_rating<-unlist(rtg)[(str_detect(unlist(rtg),"\\b\\d{3,4}P?\\b"))&(!str_detect(unlist(rtg),"\\>.?\\b\\d{3,4}P?\\b"))]
pre<-str_replace_all(pre_rating,"P","")
post_rating<-unlist(rtg)[str_detect(unlist(rtg),"\\>.?\\b\\d{3,4}P?\\b")]
post<-str_replace_all(post_rating,"([>P])","")
head(pre_rating)

```



```{r}
#Extract games played by player
games_data<-str_replace_all(str_extract_all(str_extract_all(data,"[WDL]...\\d{1,2}"),"\\.?\\d{1,2}"),"\\b[0]\\b",".")
games<-str_detect(games_data,fixed("."))
games_data<-games_data[!(games)]
head(games_data)

id<-seq(1,64,by=1)
names<-str_trim(names,"both")
state<-str_trim(state,"both")
pre_rating<-str_trim(pre_rating,"both")
post_rating<-str_trim(post_rating,"both")
player_ranks<-as.data.frame(cbind(id,names,state,pts,pre_rating,post_rating))
```

```{r}
#Calculate the mean of opponent rankings
results<-array(0,dim=nrow(player_ranks))

player_ranks[] <- lapply(player_ranks, gsub, pattern='>', replacement='')
player_ranks[] <- lapply(player_ranks, gsub, pattern='P', replacement='')


player_ranks$pts<-as.numeric(as.character(player_ranks$pts))
player_ranks$pre_rating<-as.numeric(as.character(player_ranks$pre_rating))
player_ranks$post_rating<-as.numeric(as.character(player_ranks$post_rating))

#loop through all the players
for (i in 1:nrow(player_ranks)){ 
  #extract opponent data and query rankings
  match_res<-as.numeric(str_split(unlist(str_extract_all(games_data[i],"\\d{1,2}"))," "))
  #calcuate the average of opponent rankings
results[i]<-mean(player_ranks[match_res,colnames(player_ranks)=="pre_rating"])
};

player_ranks$avg_rating<-results
head(player_ranks)

```



```{r}

# Export csv file.

write.csv(player_ranks,"player_rank",row.names=TRUE)

```



